---
title: "Text and regular expressions"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Write the R code to answer the following questions. You have until the beginning of next class to answer all of the questions below and commit to GitHub.  **It's okay if you want to do this in a .R script.  Because the data is so large, the code might run slowly, and you might not want to knit.**

# Overview

We will continue using the 'tweets' data from class containing mayor Tweets.  As a reminder, these are data shared with me.  Please do not use beyond class without inquiring with me further, and do not post publicly.

# Question 1

- Subset down to all of the tweets from Mayor Lyda Krewson (STL Mayor 2017 – 2021)
- Find the mean number of words in her tweets. Find the total number of unique words.

```{r}
# read in data
library(tidyverse)
tweets <- read_csv("https://www.dropbox.com/s/rnaaglg5eaeou6s/Tweets.csv?dl=1")

# figure out her handle
any(grepl("krewson", tweets$ScreenName))
unique(tweets$ScreenName[grepl("krewson", tweets$ScreenName)])

# subset to her tweets
lk <- tweets[tweets$ScreenName == "lydakrewson", ]

# split at spaces
split_txt <- str_split(lk$Text, pattern = " ")

# count length of each text in list
txt_length <- lapply(split_txt, length)

# unlist, then find mean of the lengths
mean(unlist(txt_length), na.rm = T)
```



# Question 2

- Come up with your own approach for searching for a dictionary of words to find Tweets relevant to some concept, such as policing, schooling, climate change, gun violence, etc.  It's up to you.  This isn't a dissertation project, so just try to look for a concept using our new regex skills.  It won't be perfect measurement, and that's totally fine.

- Then, for each mayor in the dataset (N=572), what number of tweets match your criteria?  Or, you might look at the proportion that match instead. Plot a histogram of the results using the `hist` function.

```{r}
library(stringr)
dict <- c("school", "educ", "student", "teacher", "class")
dict_regex <- paste0(dict, collapse = "|")
dict_regex <- paste0("(?i)", dict_regex)
dict_regex
```

```{r}
# Look at my dictionary with some test strings
test_strings <- c("Schooling is important!", "Our kids need to get a good education.")
str_view_all(test_strings, pattern = dict_regex)
```

```{r}
# First do it with Krewson
out <- str_subset(lk$Text, pattern = dict_regex)
length(out)
head(out)

# Create an empty dataframe for storage
# It includes the unique screenname for each mayor
educ_df <- data.frame(mayors = unique(tweets$ScreenName),
                      n_tweets = NA,
                      n_educ_tweets = NA)

# Loop over each mayor
for(i in 1:nrow(educ_df)){
  # df of just mayor's tweets
  m_tweets <- tweets$Text[tweets$ScreenName == educ_df$mayors[i]]
  # store how many total
  educ_df$n_tweets[i] <- length(m_tweets)
  # store how many match my regex
  educ_df$n_educ_tweets[i] <- length(str_subset(m_tweets, pattern = dict_regex))
}

# Plot results
hist(educ_df$n_educ_tweets/educ_df$n_tweets,
     main = "Proportion of Education Tweets",
     xlab = "")

# Who is that outlier?
which(educ_df$n_educ_tweets/educ_df$n_tweets > .25)
educ_df[139,]
```


