---
title: "Pretest"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\textbf{This pretest is designed for me to better to understand your prior knowledge of R and statistical computing. There is no harm in leaving a question blank if you do not know the answer.  This does not count toward your grade}

Please put your name at the top of this sheet of paper. You have 10 minutes to answer all of the questions below. The pretest is openbook -- you may use R, Google, etc.  Please write your answers on the paper below.

1\. Write code to calculate the mean of a vector, `x`, using only standard arithmetic operations and the `sum()` function and any other functions except `mean()`.

```{r, eval=F}
sum(x)/length(x)
```



2\. I want to make a function that will take a numeric vector, `xvec` as input, and add 2 to every element, and return the updated vector. Write in the necessary code to complete the function below.
```{r eval=F}
myFunction <- function(x){
  x+2
}
```



3\. Write out \textit{three} ways to access the first element of \texttt{myList}.
```{r}
myList <- list(letters = c("a", "B", "C"), numbers = c(1,2,3))
myList[[1]]
myList[["letters"]]
myList[[-2]]
```




4\. Write in words what the following R code accomplishes. You can assume that both
inputs are vectors of the same length.
```{r eval=FALSE}
thisFun <- function(vec1, vec2){
  output <- FALSE
    for(i in 1:length(vec1)){
      if(vec1[i] == vec2[i]){
        output <- TRUE
      }
    }
  return(output)
}
```

`thisFun()` returns `TRUE` if any element in the same position in `vec1` and `vec2`, say the fourth element, is the same.




5\. Identify the flaw in the following code:
```{r eval=FALSE}
x <- 1
while(x <= 20){
x <- (1 - x/20)
}
```

The `while()` loop will never exit. `x` will always be smaller than 1 for infinite iterations.



6\. What would the code below accomplish, assuming that the file \texttt{Simulations.csv} already exists?
```{r eval=FALSE}
cat("\n", file = "~/Simulations.csv", append = TRUE)
```

It would append an empty row to the end of the file.

7\. What function can I use to add the point $(2,1)$ to an existing plot?

The `points()` function.


