---
title: "joins"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Write the R code to answer the following questions. You have until the beginning of next class to answer all of the questions below and commit to GitHub. **You do not have to knit because of how large the data is.**

# Overview

We will continue using the Mayor Tweets data from class.  As a reminder, these are data shared with me.  Please do not use beyond class without inquiring with me further, and do not post publicly.

I've provided a third dataset of Tweets that mention the mayors' Twitter handles.  It is already cleaned so the mayor that's being mentioning is stored in the `MayorHandle` column.


```{r}
library(tidyverse)
library(dplyr)

tweets <- read_csv("https://www.dropbox.com/s/rnaaglg5eaeou6s/Tweets.csv?dl=1")
mayors <- read_csv(file = "https://raw.githubusercontent.com/jmontgomery/jmontgomery.github.io/master/PDS/Datasets/Mayors.csv")
mentions <- read_csv(file = "https://raw.githubusercontent.com/jmontgomery/jmontgomery.github.io/master/PDS/Datasets/TwitterMentions.csv")

# Some cleaning
tweets$TweetID <- as.character(tweets$TweetID)
tweets <- rename(tweets, TwitterHandle = ScreenName)

# Remove duplicates
mayors <- mayors %>%
  filter(!MayorID %in% c(1008, 1409))
```


# Question 1

Does it seem like mayors that tweet more are also mentioned more?

Erin answer:

```{r}
mentions <- rename(mentions, TwitterHandle = MayorHandle)

mention_counts <- mentions %>%
  group_by(TwitterHandle) %>%
  summarise(n_mentions = n())
mention_counts

tweet_counts <- tweets %>%
  group_by(TwitterHandle) %>%
  summarise(n_tweets = n())
tweet_counts

tweet_counts <- tweet_counts %>% full_join(mention_counts)
plot(tweet_counts$n_tweets, tweet_counts$n_mentions)
cor(tweet_counts$n_tweets, tweet_counts$n_mentions, use = "complete.obs")
```


