---
title: "dplyr"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Write the R code to answer the following questions. You have until the beginning of next class to answer all of the questions below and commit to GitHub.  **It's okay if you want to do this in a .R script.  Because the data is so large, the code might run slowly, and you might not want to knit.**

# Overview

We will continue using the polls data from class containing presidential primary polls for the 2020 election.  As a reminder, these are data shared with me.  Please do not use beyond class without inquiring with me further, and do not post publicly.

# Question 1

- Filter the data so it includes only polls taken in Jan or Feb of 2019
- Select down to only the start_date, pct, state, and candidate
- Create a new variable that is the proportion of respondents in favor
- Find the median level of support by state limited only to candidate/state combinations with at least 5 polls

```{r}
library(dplyr)
library(tidyr)
library(readr)
primaryPolls <- read_csv('president_primary_polls_feb2020.csv')
primaryPolls$start_date <- as.Date(primaryPolls$start_date, "%m/%d/%y")

# code here
```

